### 2013-07-18 doc/dissertation/document/chapters/mc_background/mc_background.tex
updating section numbering
adding a summary to each chapter

\subsubsection{Alternative Fixed Point Iterations}
\label{subsubsec:alternative_fixed_point}
In addition to the basic Richardson iteration, the MCSA algorithm
presented in Eq~(\ref{eq:mcsa}) can be used to accelerate any fixed
point iteration which only depends on the state (typically the
residual) of the last iteration. As outlined in
Appendix~\ref{ch:linear_problem}, subspace methods take on a general
form using the Petrov-Galerkin conditions as constraints for
extracting a correction from the search subspace as given by
Eq~(\ref{eq:linear_projection_iteration}). Those that are
one-dimensional may be used with fixed-point MCSA as they only depend
the previous iteration for information. As an example, for
positive-definite but not necessarily symmetric problems, the minimal
residual (MINRES) iteration \citep{saad_iterative_2003} can be used in
conjunction with MCSA where the residual vector, $\mathbf{r}$, defines
the search subspace and the action of the linear operator on the
residual vector, $\mathbf{A}\mathbf{r}$, defines the constraint
subspace. Using this, we can then define an MCSA scheme that
accelerates the MINRES iteration:
\begin{subequations}
  \begin{gather}
    \alpha = \frac{\langle \mathbf{A}\mathbf{r}^k, \mathbf{r}^k
      \rangle}{\langle \mathbf{A}\mathbf{r}^k, \mathbf{A}\mathbf{r}^k
      \rangle}\:, \\
    \ve{x}^{k+1/2} = \ve{x}^k + \alpha \ve{r}^k\:,
    \label{eq:minres_correction}\\
    \ve{r}^{k+1/2} = \ve{b} - \ve{A}\ve{x}^{k+1/2}\:,\\ 
    \ve{A}\delta\ve{x}^{k+1/2} = \ve{r}^{k+1/2}\:,\\ 
    \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta\ve{x}^{k+1/2}\:,\\ 
    \ve{r}^{k+1} = \ve{b} - \ve{A}\ve{x}^{k+1}\:.
  \end{gather}
  \label{eq:mcsa_minres}
\end{subequations}
where $\alpha$ is an optimal extrapolation parameter generated from
the constraints. Interestingly, this scheme is nearly identical to the
Richardson iteration version in Eq~(\ref{eq:mcsa}) except now the
additional extrapolation parameter, $\alpha$, is applied to the
residual correction in Eq~(\ref{eq:minres_correction}) as a means of
selecting a more optimal search direction at each iteration,
potentially further accelerating convergence.

@@

>>>>

\subsection{Alternative Fixed Point Iterations}
\label{subsubsec:alternative_fixed_point}
In addition to the basic Richardson iteration, the MCSA algorithm
presented in Eq~(\ref{eq:mcsa}) can be used to accelerate any fixed
point iteration which only depends on the state (typically the
residual) of the last iteration. As outlined in
Appendix~\ref{ch:linear_problem}, subspace methods take on a general
form using the Petrov-Galerkin conditions as constraints for
extracting a correction from the search subspace as given by
Eq~(\ref{eq:linear_projection_iteration}). Those that are
one-dimensional may be used with fixed-point MCSA as they only depend
the previous iteration for information. As an example, for
positive-definite but not necessarily symmetric problems, the minimal
residual (MINRES) iteration \citep{saad_iterative_2003} can be used in
conjunction with MCSA where the residual vector, $\mathbf{r}$, defines
the search subspace and the action of the linear operator on the
residual vector, $\mathbf{A}\mathbf{r}$, defines the constraint
subspace. Using this, we can then define an MCSA scheme that
accelerates the MINRES iteration:
\begin{subequations}
  \begin{gather}
    \alpha = \frac{\langle \mathbf{A}\mathbf{r}^k, \mathbf{r}^k
      \rangle}{\langle \mathbf{A}\mathbf{r}^k, \mathbf{A}\mathbf{r}^k
      \rangle}\:, \\
    \ve{x}^{k+1/2} = \ve{x}^k + \alpha \ve{r}^k\:,
    \label{eq:minres_correction}\\
    \ve{r}^{k+1/2} = \ve{b} - \ve{A}\ve{x}^{k+1/2}\:,\\ 
    \ve{A}\delta\ve{x}^{k+1/2} = \ve{r}^{k+1/2}\:,\\ 
    \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta\ve{x}^{k+1/2}\:,\\ 
    \ve{r}^{k+1} = \ve{b} - \ve{A}\ve{x}^{k+1}\:.
  \end{gather}
  \label{eq:mcsa_minres}
\end{subequations}
where $\alpha$ is an optimal extrapolation parameter generated from
the constraints. Interestingly, this scheme is nearly identical to the
Richardson iteration version in Eq~(\ref{eq:mcsa}) except now the
additional extrapolation parameter, $\alpha$, is applied to the
residual correction in Eq~(\ref{eq:minres_correction}) as a means of
selecting a more optimal search direction at each iteration,
potentially further accelerating convergence.

@@

%%---------------------------------------------------------------------------%%
\section{Summary}
\label{sec:mc_summary}

In this chapter, the fundamentals of Monte Carlo synthetic
acceleration methods were presented and explored in the context of a
simple model transport system. The following are the signficant
observations and findings.

\begin{itemize}
\item MCSA has been generalized for all discrete linear transport
  problems
\item MCSA can be preconditioned using a general left/right scheme but
  requires the explicit inversion of the preconditioners and
  construction of the composite linear operator
\item The adjoint Neumann-Ulam method performs better when used with
  MCSA compared to the forward method
\item MCSA has superior performance to Halton's residual Monte Carlo
  method
\item The Richardson iteration in MCSA behaves as a smoother for the
  stochastic Monte Carlo error
\item MCSA iterative performance when used with the expected value
  estimator is insensitive to histories per iteration
\item MCSA iterative performance is insensitive to the stochastic
  history weight cutoff
\item Variance reduction with artificial absorption can enhance time
  to solution
\item The reduced domain approximation was introduced to mitigate the
  consequences of composite linear operator generation
\end{itemize}

### 2013-07-18 doc/dissertation/document/chapters/nonlinear_problem/nonlinear_problem.tex
adding a summary to each chapter

\chapter{Monte Carlo Synthetic Acceleration Methods for the
  Navier-Stokes \\Equations\ }
\label{ch:nonlinear_problem}
Nonlinear transport problems are a common occurrence in single and
multiple physics problems in nuclear engineering. Systems of partial
differential equations, such as the Navier-Stokes equations that
describe fluid flow, yield discrete sets of stiff equations with
nonlinearities present in the variables when discretized by
conventional methods. These sets of equations characterize momentum
and energy transport through diffusion, viscous forces, and inertial
forces. Traditionally, such systems have been solved by linearizing
them in a form where the nonlinearities in the variables are
eliminated and more traditional linear methods can be used for
solutions. Often characterized as segregated methods where physics
operators are split and their action on the system approximated in
steps, such methods lack consistency and accuracy in resolving the
nonlinear component of the solution. In the last 30 years, fully
consistent nonlinear methods based on Newton's method have become more
popular and many advances have been made in the nuclear engineering
field to employ these methods.

@@

>>>>

\chapter{Monte Carlo Synthetic Acceleration Methods for the
  Navier-Stokes \\Equations\ }
\label{ch:nonlinear_problem}
Nonlinear transport problems are a common occurrence in single and
multiple physics problems in nuclear engineering. Systems of partial
differential equations, such as the Navier-Stokes equations that
describe fluid flow, yield discrete sets of stiff equations with
nonlinearities present in the variables when discretized by
conventional methods. These sets of equations characterize momentum
and energy transport through diffusion, viscous forces and convection,
and inertial forces. Traditionally, such systems have been solved by
linearizing them in a form where the nonlinearities in the variables
are eliminated and more traditional linear methods can be used for
solutions. Often characterized as segregated methods where physics
operators are split and their action on the system approximated in
steps, such methods lack consistency and accuracy in resolving the
nonlinear component of the solution. In the last 30 years, fully
consistent nonlinear methods based on Newton's method have become more
popular and many advances have been made in the nuclear engineering
field to employ these methods.

@@

%%---------------------------------------------------------------------------%%
\section{Summary}
\label{sec:nonlinear_summary}

In this chapter we have developed and explored Monte Carlo synthetic
acceleration methods in the context of solutions to the nonlinear
Navier-Stokes equations. The following are the signficant observations
and findings.

\begin{itemize}
\item Forward-Automated Newton-MCSA (FANM), a new inexact Newton
  method based on Monte Carlo synthetic acceleration, has been
  developed
\item The FANM method has been incorporated into the Drekar production
  multiphysics code
\item The FANM method has been verified to produce the same solutions
  as a production Newton-Krylov method for two benchmark problems for
  the Navier-Stokes equations in different flow regimes
\item The FANM method has better iterative performance than the
  Newton-Krylov method for convection dominated problems, converging
  in fewer linear solver iterations with the same preconditioning for
  high and lower Rayleigh numbers
\item The Newton-Krylov method has better iterative performance than
  the FANM method for flow dominated by interial forces, converging in
  fewer linear solver iterations with the same preconditioning
\item The spectral radius convergence restriction on MCSA was observed
  to be a signficant hinderance by preventing solutions to forced flow
  problems at high Reynolds numbers
\item Explict algebraic preconditioning of MCSA creates FANM run times
  $O(100)$ slower than the Newton-Krylov solver
\item More Monte Carlo histories at every FANM iteration can reduce
  the number of linear and nonlinear iterations required to converge
  the problem
\end{itemize}

### 2013-07-18 doc/dissertation/document/chapters/parallel_mc/parallel_mc.tex
adding a summary to each chapter

>>>>

%%---------------------------------------------------------------------------%%
\section{Summary}
\label{sec:parallel_summary}

In this chapter we have developed a domain-decomposed parallel
algorithm for Monte Carlo synthetic acceleration and explored the
qualities of the algorithm through parallel scaling studies. The
following are the signficant observations and findings.

\begin{itemize}
\item MCSA can be effectively parallelized using domain decomposition
\item The leakage fraction of stochastic histories from a parallel
  subdomain can be quantified analytically using algebraic quantities
\item The multiple-set overlapping-domain (MSOD) parallel algorithm
  for particle transport has been adapted to parallelize MCSA
\item MCSA parallelized with MSOD was verified to produce the same
  results for solutions to the neutron diffusion equation as
  production Krylov methods
\item Strong scaling studies were performed with up to 65,536 cores
  and weak scaling studies were performed with up to 131,072 cores
\item MCSA scales comparably to production Krylov methods on the Titan
  Cray XK7 leadership class machine
\item Overlap in small quantities determined analytically by random
  walk length can provide parallel efficiency boosts of a few percent
  in strong scaling cases but is ineffective in weak scaling cases
\item Multiple sets will never boost parallel efficienicies in both
  weak and strong scaling cases due to additional communication
  overhead
\item Multiple sets offers a means to reduce time to solution by
  solving a multiple copies of the original problem and combining the
  solutions using superposition
\item MCSA is most efficient used in parallel as a stochastic
  realization of an additive Schwarz method
\end{itemize}

### 2013-07-18 doc/dissertation/document/chapters/spn_equations/spn_equations.tex
adding a summary to each chapter

>>>>

%%---------------------------------------------------------------------------%%
\section{Summary}
\label{sec:mc_summary}

In this chapter, the Monte Carlo synthetic acceleration method has
been applied to the $SP_N$ discretization of the neutron transport
equation and explored in the context of a difficult nuclear fuel
assembly criticality problem. The following are the signficant
observations and findings.

\begin{itemize}
\item MCSA can be used to solve the asymmetric system generated by the
  multigroup $SP_N$ equations
\item Simple Jacobi-based preconditioning reduces the spectral radius
  of the $SP_N$ system below unity, however, this is not sufficient
  for MCSA convergence
\item As the spectral radius approaches unity, MCSA was demonstrated
  to break down with more stochastic histories required for
  convergence and more CPU time required per history
\item Advanced algebraic preconditioning strategies were applied to
  the $SP_N$ equations with ILUT chosen for subsequent investigations
\item The reduced domain approximation was required to mitigate the
  memory constraints of the explicit preconditioning strategy
\item MCSA relaxation parameters were developed an observed to enhance
  the time to solution for the fuel assembly problem
\item MCSA was verified to produce the same flux distribution and
  k-eigenvalue for the fuel assembly as production Krylov methods
\item MCSA was observed to converge in fewer iterations per eigenvalue
  iteration than GMRES for the fuel assembly problem and more then
  Bi-CGStab
\item The explicit preconditioning strategy creates MCSA runtimes
  $O(100)$ larger than the production Krylov methods
\end{itemize}
